% !TEX root = dsge-model-appendix.tex
\input{DynareModWizard/appendix/__header.tex}

%%%%%%%%%%%%%%%%%%%%%%%
% MODEL SPECIFICATION %
%%%%%%%%%%%%%%%%%%%%%%%
\setboolean{openEconomy}{true}
\renewcommand{\currencyParadigm}{PCP}
\setboolean{TANK}{false}
\setboolean{capitalUtilization}{false}
\setboolean{habitFormation}{true}
\setboolean{utilityEnhancingGovernmentSpending}{false}
\setboolean{utilityEnhancingBonds}{false}
\setboolean{importAdjustmentCosts}{true}
\setboolean{governmentBonds}{true}
\setboolean{taxCapital}{false}
\setboolean{taxConsumption}{false}
\setboolean{taxLabor}{false}
\setboolean{icebergTradeCosts}{true}
\setboolean{intermediateInputsInProduction}{true}
\setboolean{importContentGovernmentSpending}{false}

\begin{document}

\title{The Price of War}
\subtitle{DSGE model documentation}
\author{Jonathan Federle, 
André Meier,
Gernot J. Müller,\\
Willi Mutschler,
Moritz Schularick}
\date{May 2024\\~\\NOT FOR PUBLICATION}

\maketitle%
\begin{abstract}
\noindent
This document derives in detail the multi-country DSGE model used in section 4 of the paper \emph{The Price of War}.
Replication codes are written in Dynare and available at
\begin{center}\url{https://github.com/wmutschl/price-of-war}\end{center}
Please raise any issues or questions as an issue on GitHub (or get in touch via email).
\end{abstract}

\newpage
\tableofcontents
\newpage

\section{Overview}
\begin{itemize}[itemsep=0pt, topsep=0pt, partopsep=0pt, parsep=0pt]
\item Dynamic Stochastic General Equilibrium (DSGE) model with multiple countries denoted by Home (\(H\)), Nearby (\(N\)), Distant (\(D\)),
  and Rest of World (\(R\)).
\item Focus is on the impact of a war shock in Home and the international spillovers.
\end{itemize}

\paragraph{Households}
\begin{itemize}[itemsep=0pt, topsep=0pt, partopsep=0pt, parsep=0pt]
  \item Households maximize lifetime utility considering consumption, labor supply and habit formation.
  \item Capital accumulation is governed by a law of motion with investment adjustment costs.
  \item Monopolistic competitive households set wage for distinct labor types according to Calvo frictions.
  \item Households hold foreign bonds and use imported goods for domestic absorption.
  \item Budget constraint includes income from labor, capital, bonds and government transfers.  
\end{itemize}

\paragraph{Firms}
\begin{itemize}[itemsep=0pt, topsep=0pt, partopsep=0pt, parsep=0pt]
  \item Retail goods are distributed using a CES aggregation of locally produced and imported goods, subject to import adjustment costs and iceberg trade frictions.
  \item Wholesale goods are produced by monopolistic competitive firms using labor, capital and intermediate inputs in production.
  Price setting is subject to Calvo price frictions.
  \item Export prices are set under the producer currency pricing (PCP) regime.
\end{itemize}

\paragraph{Government}
\begin{itemize}[itemsep=0pt, topsep=0pt, partopsep=0pt, parsep=0pt]
  \item Monetary Policy is defined by an interest rate rule with feedback to inflation deviations and output gap.
  \item Fiscal authority finances government spending by issuing bonds and lump-sum taxes.
\end{itemize}

\paragraph{Market clearing}
\begin{itemize}[itemsep=0pt, topsep=0pt, partopsep=0pt, parsep=0pt]
  \item Sequence-of-markets concept ensures equilibrium in capital, labor, bond and goods markets.
  \item Price and wage inefficiencies are captured by dispersion terms.
  \item Test of market clearing by including all aggregated budget constraints (Walras law).
\end{itemize}

\paragraph{War shock}
\begin{itemize}[itemsep=0pt, topsep=0pt, partopsep=0pt, parsep=0pt]
  \item Autoregressive process of order two.
  \item Home's productivity declines, trade costs with Home go up and capital is destroyed in Home.
  \item All countries increase government spending to finance military expenditures, with varying magnitudes.
  \item Section~\ref{sec:WarShock} provides details which equations are affected by the war shock.  
\end{itemize}

\newpage

\paragraph{Implementation in Dynare}
\begin{itemize}[itemsep=0pt, topsep=0pt, partopsep=0pt, parsep=0pt]
  \item Model equations are derived in per-capita variables.
  \item Gray boxes contain model equations as entered in Dynare.
\end{itemize}

\paragraph{Related literature and references}
The model and its derivation are inspired by the literature on open-economy DSGE models
  and we draw heavily on codes and technical appendices of the following references:
\begin{itemize}
\item \textcite{Cwik.Muller.Wolters_2011_DoesTradeIntegration}
\item \textcite{Eichenbaum.Johannsen.Rebelo_2021_MonetaryPolicyPredictability}
\item \textcite{Erceg.Guerrieri.Gust_2006_SIGMANewOpen}
\item \textcite{Georgiadis.Schumann_2021_DominantcurrencyPricingGlobal}
\item \textcite{Gomes.Jacquinot.Pisani_2012_EAGLEModelPolicy}
\item \textcite{Gopinath.Boz.Casas.EtAl_2020_DominantCurrencyParadigm}
\end{itemize}
\newpage

\section{Notation}
\input{DynareModWizard/appendix/_notation.tex}
\newpage

\section{Households}
\input{DynareModWizard/appendix/_households.tex}

\subsection{Labor supply and wage setting\label{sec:LaborSupply}}
\input{DynareModWizard/appendix/_households_labor_calvo.tex}

\section{Firms}

\subsection{Retail goods}
\input{DynareModWizard/appendix/_firms_retail_distribution.tex}

\subsection{Wholesale goods}
\ifthenelse{\equal{\currencyParadigm}{mixed}}{%
  \subsubsection{Import distribution}
}{}%
\input{DynareModWizard/appendix/_firms_wholesale_distribution_imports.tex}

\ifthenelse{\equal{\currencyParadigm}{LCP}\OR\equal{\currencyParadigm}{mixed}}{%
\subsubsection{Export good aggregation (LCP)}
\renewcommand{\XCP}{LCP}
\input{DynareModWizard/appendix/_firms_wholesale_aggregation.tex}
}{}%

\ifthenelse{\equal{\currencyParadigm}{PCP}\OR\equal{\currencyParadigm}{mixed}}{%
\subsubsection{Export good aggregation (PCP)}
\renewcommand{\XCP}{PCP}
\input{DynareModWizard/appendix/_firms_wholesale_aggregation.tex}
}{}%

\ifthenelse{\equal{\currencyParadigm}{DCP}\OR\equal{\currencyParadigm}{mixed}}{%
\subsubsection{Export good aggregation (DCP)}
\renewcommand{\XCP}{DCP}
\input{DynareModWizard/appendix/_firms_wholesale_aggregation.tex}
}{}%

\subsubsection{Local good aggregation}
\renewcommand{\XCP}{local}
\input{DynareModWizard/appendix/_firms_wholesale_aggregation.tex}

\subsection{Intermediate goods}
\input{DynareModWizard/appendix/_firms_intermediate.tex}

\ifthenelse{\equal{\currencyParadigm}{LCP}\OR\equal{\currencyParadigm}{mixed}}{%
\renewcommand{\XCP}{LCP}
\subsubsection{Export pricing (\XCP)}
\input{DynareModWizard/appendix/_firms_intermediate_monopolists_pricing_calvo.tex}
}{}%

\ifthenelse{\equal{\currencyParadigm}{PCP}\OR\equal{\currencyParadigm}{mixed}}{%
\renewcommand{\XCP}{PCP}
\subsubsection{Export pricing (\XCP)}
\input{DynareModWizard/appendix/_firms_intermediate_monopolists_pricing_calvo.tex}
}{}%

\ifthenelse{\equal{\currencyParadigm}{DCP}\OR\equal{\currencyParadigm}{mixed}}{%
\renewcommand{\XCP}{DCP}
\subsubsection{Export pricing (\XCP)}
\input{DynareModWizard/appendix/_firms_intermediate_monopolists_pricing_calvo.tex}
}{}%

\renewcommand{\XCP}{local}
\subsubsection{Local pricing}
\input{DynareModWizard/appendix/_firms_intermediate_monopolists_pricing_calvo.tex}

\section{Government}

\subsection{Monetary policy}
\input{DynareModWizard/appendix/_policy_monetary.tex}

\subsection{Fiscal policy}
\input{DynareModWizard/appendix/_policy_fiscal.tex}

\section{Market clearing}

\subsection{Capital market}
\input{DynareModWizard/appendix/_clearing_market_capital.tex}

\subsection{Labor market}
\input{DynareModWizard/appendix/_clearing_market_labor.tex}

\subsection{Aggregation}
\input{DynareModWizard/appendix/_clearing_market_aggregation.tex}

\subsection{Bond market}
\input{DynareModWizard/appendix/_clearing_market_bonds.tex}

\subsection{Aggregate demand}
\input{DynareModWizard/appendix/_clearing_market_goods_demand.tex}

\subsection{Aggregate supply}
\input{DynareModWizard/appendix/_clearing_market_goods_supply.tex}

\subsection{Aggregate budget constraint}
\input{DynareModWizard/appendix/_clearing_market_budget.tex}

\newpage
\section{War shock scenario}
The Home economy is impacted by a war shock, which is modeled as an autoregressive process.
This shock triggers a sequence of events influencing the economy's dynamics:
  trade costs rise due to increased iceberg costs with Home,
  capital in Home is destroyed,
  and Home's productivity declines.
Moreover, all governments increase their spending to finance additional military and defense expenditures.

\paragraph{War shock\label{sec:WarShock}}
The war shock, denoted by \(\omega_{t}\), is modeled to follow an autoregressive process of order two
  (AR{(2)}) with persistence parameters \(\rho_{1}\) and \(\rho_{2}\) and an exogenous innovation \(\varepsilon^{\omega}_{t}\),
  which triggers the war shock by assuming a value of 1 at the onset of war and 0 otherwise.
\begin{tcolorbox}[colback=lightGray, boxrule=1pt, center title, title=\textbf{War Shock}]
war shock process
\begin{align*}
\omega_{t} = \rho^{\omega}_{1} \omega_{t-1} + \rho^{\omega}_{2} \omega_{t-2} + \varepsilon^{\omega}_{t}
\end{align*}
\end{tcolorbox}
To ensure the process remains stationary, the roots of the war shock process,
  denoted by \(\rho^{\omega}_{I}\) and \(\rho^{\omega}_{II}\),
  are constrained to lie within the unit circle:
\begin{align*}
\rho^{\omega}_{1} = \rho^{\omega}_{I} + \rho^{\omega}_{II}, && \rho^{\omega}_{2} = - \rho^{\omega}_{I} \cdot \rho^{\omega}_{II}
\end{align*}

\paragraph{Iceberg trade costs with Home}
The war shock increases iceberg trade costs of all countries
  in their trade relationship with Home
  according to the war shock scaled by \(\Delta^{\mathbb{o}}\geq0\).
\begin{tcolorbox}[colback=lightGray, boxrule=1pt, center title, title=\textbf{Iceberg Trade Costs}]
iceberg trade costs in warring trade relationships (\(i\neq H\))
\begin{align*}
\tau^{\mathbb{o}}_{iH,t} = \tau^{\mathbb{o}}_{Hi,t} = 1 + \Delta^{\mathbb{o}} \omega_{t}
\end{align*}
no iceberg trade costs in non-warring trade relationships (\(j \neq H\)):
\begin{align*}
\tau^{\mathbb{o}}_{ij,t} = \tau^{\mathbb{o}}_{ji,t} = 1
\end{align*}
\end{tcolorbox}

\paragraph{Total factor productivity reduction in Home}
The war shock impacts total factor productivity in Home, \(A_{H,t}\), by reducing it according to the war shock scaled by \(\Delta^{A}_{H}\).
\begin{tcolorbox}[colback=lightGray, boxrule=1pt, center title, title=\textbf{Total Factor Productivity Reduction}]
total factor productivity in \(H\):
\begin{align*}
\log(A_{H,t}) = (1-\rho^{A}_{H}) \log(\overline{A}_{H}) + \rho^{A}_{H} \log(A_{H,t-1}) - \Delta^{A}_{H} \omega_{t}
\end{align*}
\end{tcolorbox}

\paragraph{Capital destruction in Home}
The war shock destroys capital in Home, \(K_{H,t}\), scaled by a factor \(\Delta^{K}_{H}\).
This modification changes the capital accumulation equation
  and, accordingly, the optimal capital and investment decisions in Home.
\begin{tcolorbox}[colback=lightGray, boxrule=1pt, center title, title=\textbf{Capital Destruction}]
capital accumulation (per-capita) in \(H\):
\begin{align*}
k_{H,t} = \left(
  (1-\delta^{K}_{H}) k_{H,t-1}
+ \left( 1 - \frac{\phi^{K}_{H}}{2} {\left(\frac{i_{H,t}}{i_{H,t-1}} - 1 \right)}^2 \right) i_{H,t}
\right) e^{-\Delta^{K}_{H} \omega_{t}}
\end{align*}
optimal capital decision (per-capita) in \(H\):
\begin{align*}
\lambda_{H,t} q^{K}_{H,t} = \beta_{H} \mathbb{E}_{t} \lambda_{H,t+1} \left( q^{K}_{H,t+1}(1-\delta^{K}_{H}) e^{-\Delta^{K}_{H} \omega_{t+1}} + r^{K}_{H,t+1} \right)
\end{align*}
optimal investment decision (per-capita) in \(H\):
\begin{multline*}
\lambda_{H,t} = e^{-\Delta^{K}_{H} \omega_{t}} \lambda_{H,t} q^{K}_{H,t}
\Bigg (
  1 - \frac{\phi^{K}_{H}}{2} {\left(\frac{i_{H,t}}{i_{H,t-1}}-1\right)}^2
  - \phi^{K}_{H} \left( \frac{i_{H,t}}{i_{H,t-1}} - 1 \right) \left(\frac{i_{H,t}}{i_{H,t-1}}\right)
\Bigg )
\\
+ \beta_{H} \mathbb{E}_{t} e^{-\Delta^{K}_{H} \omega_{t+1}} \lambda_{H,t+1} q^{K}_{H,t+1}
    \phi^{K}_{H} \left( \frac{i_{H,t+1}}{i_{H,t}} - 1\right)
    {\left(\frac{i_{H,t+1}}{i_{H,t}}\right)}^2
\end{multline*}
\end{tcolorbox}

\paragraph{Military spending in all countries}
All countries increase their government spending to finance additional military expenditures;
  however, the magnitude of the increase differs between Home and the other countries.
\begin{tcolorbox}[colback=lightGray, boxrule=1pt, center title, title=\textbf{Military Spending}]
government spending in \(H\)
\begin{align*}
\log\left(\frac{g_{H,t}}{g_{H}}\right) = \rho^{G}_{H} \log\left(\frac{g_{H,t-1}}{g_{H}}\right) + \Delta^{G}_{H} \omega_{t}
\end{align*}
government spending in \(j \neq H\)
\begin{align*}
\log\left(\frac{g_{j,t}}{g_{j}}\right) = \rho^{G} \log\left(\frac{g_{j,t-1}}{g_{j}}\right) + \Delta^{G} \omega_{t}
\end{align*}
\end{tcolorbox}

\newpage
\printbibliography%

\end{document}